---
title: "Data Visualization Code"
output: html_document
date: "2024-05-06"
---
This shows the code to visualize bandpower values across MEG gradiometers, from the MEG data collection of one articulated speech sample and 1 imagined speech sample. This is based off of the data and code I used for the 1st steps of my data replication project, but with some changes to better suit it to the Shiny Web App visualization



Loading in the Data
```{r, echo=FALSE}
library(tidyverse)
f_ar <- "data/ar1.csv"
f_im <- "data/im1.csv"
ar1 <- read_csv(f_ar, col_names = FALSE)
im1 <- read_csv(f_im, col_names = FALSE)
```


Establishing Brainwave Band Frequency Ranges
```{r}
frequencyBands <- list(
  Delta = c(0.3, 4),
  Theta = c(4, 8),
  Alpha = c(8, 15),
  Beta = c(15, 30),
  Gamma = c(30, 59),
  Lower_HighGamma = c(61, 119),
  Upper_HighGamma = c(121, 250)
)
```


Creating a Bandpower Function
```{r}
Fs <- 1000 # a priori values for sampling frequency

# Creating a bandpower function
bandpower_funct <- function(signal, freq_low, freq_high, Fs) {
  # how many data points are in the signal?
  n <- length(signal)
  # using the built-in fft function in R
  fft_res <- fft(signal)
  # data indices should correspond to respective frequency value
  # sampling frequency, Fs, is split into, n, evenly spaced 'bins'--> (Fs/n)
  freq_val <- (0:(n-1)) * (Fs / n)
  # calculate power spectral density (PSD)
  psd <- abs(fft_res)^2 / n
  # which() to pick out indices for values that fit the criteria
  # the criteria: frequencies between the low and high bounds for the given band
  band_indices <- which(freq_val >= freq_low & freq_val <= freq_high)
  # CALCULATE BANDPOWER
  # sum as the PSD values for the matrices then divide by the number of data points
  bandpower <- sum(psd[band_indices]) / length(band_indices)
  return(bandpower)
}
```

Storing Articulation Bandpower Values
```{r}
# store bandpower values for each band into a list
art1_bandpowers <- list()

# for each iteration/band in the frequencyBands list created before
for(band in names(frequencyBands)) {
  band_low <- frequencyBands[[band]][1] # first value is lower bound for respective band
  band_high <- frequencyBands[[band]][2] # second value is upper bound
  
  # creating a matrix to store the resulting bandpower values
  # as many rows as there are sensors/rows in ar1, and 1 column --> this goes for each pass(band)
  art1_band_matrix <- matrix(nrow=nrow(ar1), ncol=1)
  
  # for each sensor pass all the way through the end of the rows(the last sensor in ar1)
  for(sensor_idx in 1:nrow(ar1)) {
    # apply the bandpower function and assign to the band_matrix
    art1_band_matrix[sensor_idx, 1] <- bandpower_funct(as.numeric(ar1[sensor_idx, ]), band_low, band_high, Fs)
  }
  
  # now store all the bandpower values
  art1_bandpowers[[band]] <- art1_band_matrix
}
```

Storing Imagination Bandpower Values
```{r}
# store bandpower values for each band into a list
im1_bandpowers <- list()

# for each iteration/band in the frequencyBands list created before
for(band in names(frequencyBands)) {
  band_low <- frequencyBands[[band]][1] # first value is lower bound for respective band
  band_high <- frequencyBands[[band]][2] # second value is upper bound
  
  # creating a matrix to store the resulting bandpower values
  # as many rows as there are sensors/rows in im1, and 1 column --> this goes for each pass(band)
  im1_band_matrix <- matrix(nrow=nrow(im1), ncol=1)
  
  # for each sensor pass all the way through the end of the rows(the last sensor in im1)
  for(sensor_idx in 1:nrow(im1)) {
    # apply the bandpower function and assign to the band_matrix
    im1_band_matrix[sensor_idx, 1] <- bandpower_funct(as.numeric(im1[sensor_idx, ]), band_low, band_high, Fs)
  }
  
  # now store all the bandpower values
  im1_bandpowers[[band]] <- im1_band_matrix
}
```


Reshaping the data into a long format- this is more suitable for ggplot
```{r}
# creating a function to convert the matrices into data frames
# input = bandpower file (art1 or im1) and filename 
prepare_bandpower_data <- function(bandpowerfile, filetitle) { 
  band_data <- map_df(names(bandpowerfile), function(band) {
    data_frame(
      Sensor = 1:nrow(bandpowerfile[[band]]),
      Bandpower = as.vector(bandpowerfile[[band]]),
      Band = band,
      File_title = filetitle
    )
  })
  return(band_data)
}

# Prepare data for both conditions
art1_data <- prepare_bandpower_data(art1_bandpowers, "Articulation")
im1_data <- prepare_bandpower_data(im1_bandpowers, "Imagination")

# combined_data has the new articulation AND imaginationd data in the long format
combined_data <- bind_rows(art1_data, im1_data)
```



```{r}
library(ggplot2)

# plot each band separately
(plots <- lapply(unique(combined_data$Band), function(band) {
  ggplot(data = combined_data %>% filter(Band == band), 
         aes(x = File_title, y = Bandpower, color = File_title)) +
    geom_point(position = position_jitterdodge(jitter.width = 0.2, dodge.width = 0.75), size = 3) +
    labs(title = paste("Bandpower Values for", band, "Band"),
         x = "",
         y = "Bandpower") +
    scale_color_manual(values = c("Articulation" = "blue", "Imagination" = "red")) +
    theme_minimal() +
    theme(legend.position = "none")
}))


```


